Probabilistic Catalogs for Crowded Stellar Fields

### 2012-11-22 paper/ms.tex
\begin{figure}[ht!]
Don't name journal
Tweaks, including preprint format (looks great!)

\documentclass[manuscript]{aastex}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citet and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}	% for blackboard bold numbers

@@

\slugcomment{To be submitted to the Astronomical Journal}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{The two simulated images used to test our methodology.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference1.eps}
\end{center}
\caption{Inference about the parameters for Test Case 1. Note that there is
considerable uncertainty (particularly about $h_2$), which occurs because the
shape of the broken power-law does not depend strongly on the parameters.
The true input values are plotted as filled squares.
\label{fig:results1}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference2.eps}
\end{center}
\caption{Inference about the parameters for Test Case 2. Note that the
parameters are still not very well constrained even with the larger number of
stars. The true input values are plotted as filled squares.\label{fig:results2}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{Figures/catalogs.eps}
\end{center}
\caption{Nine example catalogs sampled from the posterior distribution for
Test Case 2. Features in common represent features with high probability,
and differences between the catalogs represent conclusions that are uncertain.
The area of each circle is proportional to the flux of the star.\label{fig:catalogs}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images for test case 2. The upper left panel shows the posterior
mean high-resolution scene. The upper right panel shows the posterior mean scene
when observed at the resolution of the data, and the bottom panels show the
model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_N.eps}
\end{center}
\caption{The number of stars in the catalogs produced by \sex~compared with the
actual number of stars in the images. \sex~finds only a small fraction of the
stars actually present. This is to be expected in the high-$\sigma$ case as
\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
end the number of stars is still drastically underestimated. Note that the true
numbers
of stars shown in these plots are $\sim$82\% of the values listed in
Figure~\ref{tab:truth} because here we only count those stars with positions
in $[-1, 1] \times [-1, 1]$.
\label{fig:sex_N}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
\end{center}
\caption{Inference of the luminosity function parameters using the $0.5\sigma$
catalog from \sex. Note that the inference appears to rule out regions around
the true input values. This occurs because the uncertainties in the existence
and fluxes of the faint stars are not taken into account. The true input values are plotted as filled squares.
\label{fig:sex_inference}}
\end{figure}

@@

We note that there are many limitations to the model presented in this paper.
In principle, our model should be a model of the physical state of the universe,
and not a simple model where the only stellar properties are a $2-D$ position and
a flux. Another limitation is that we have not considered multi-epoch or multi-band
imaging. In the former case, PSF variations and stellar motions may be relevant
\citep{lang}, and in the latter, a model for the spectral energy distributions
of the stars will need to be considered.

@@

\section{Acknowledgements}
It is a pleasure to thank
  Jonathan Goodman (NYU),
  Fengji Hou (NYU),
  Dustin Lang (CMU),
  Geraint Lewis (Sydney), and
  Phil Marshall (Oxford)
for their comments and discussions.
BJB would also
like to thank Wayne Stewart and Arden Miller for their support during
his first semester at Auckland.
DFM and DWH were partially supported by
   the US National Science Foundation (grant IIS-1124794) and
   the US National Aeronautics and Space Administration (grant NNX12AI50G).

@@

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcl}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcrl}
0, & & x < h_1 & \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 & \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
Two conditions will be used to determine the normalizers $Z_1$ and $Z_2$.
Firstly, the probability density function (PDF) should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}
The second condition is that the total probability must be 1:
\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}
\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}.
\end{eqnarray}
The cumulative distribution (CDF) is a useful property of a probability
distribution and is given by the antiderivative of the PDF:
\begin{eqnarray}
P(X \leq x) = F(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The inverse of the CDF is also useful and is given by:
\begin{eqnarray}
F^{-1}(u) &=&
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1.
\end{array}
\right.
\end{eqnarray}
An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/broken.eps}
\caption{A broken power-law distribution. The parameter values for this
particular PDF were
$\{h_1, h_2, \alpha_1, \alpha_2\} = \{0.3, 0.6, 1.1, 2\}$, i.e. the same
parameter values used to make the simulated data.
\label{fig:powerlaw}}
\end{center}
\end{figure}

>>>>

\documentclass[12pt, preprint]{aastex}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citet and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}	% for blackboard bold numbers

@@

%\slugcomment{To be submitted}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/test_cases.eps}
\caption{The two simulated images used to test our methodology.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference1.eps}
\end{center}
\caption{Inference about the parameters for Test Case 1. Note that there is
considerable uncertainty (particularly about $h_2$), which occurs because the
shape of the broken power-law does not depend strongly on the parameters.
The true input values are plotted as filled squares.
\label{fig:results1}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/inference2.eps}
\end{center}
\caption{Inference about the parameters for Test Case 2. Note that the
parameters are still not very well constrained even with the larger number of
stars. The true input values are plotted as filled squares.\label{fig:results2}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[scale=0.6]{Figures/catalogs.eps}
\end{center}
\caption{Nine example catalogs sampled from the posterior distribution for
Test Case 2. Features in common represent features with high probability,
and differences between the catalogs represent conclusions that are uncertain.
The area of each circle is proportional to the flux of the star.\label{fig:catalogs}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/summaries.eps}
\end{center}
\caption{Summary images for test case 2. The upper left panel shows the posterior
mean high-resolution scene. The upper right panel shows the posterior mean scene
when observed at the resolution of the data, and the bottom panels show the
model residuals.
\label{fig:summaries}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_N.eps}
\end{center}
\caption{The number of stars in the catalogs produced by \sex~compared with the
actual number of stars in the images. \sex~finds only a small fraction of the
stars actually present. This is to be expected in the high-$\sigma$ case as
\sex~then only searches for stars above a certain flux, but at the low-$\sigma$
end the number of stars is still drastically underestimated. Note that the true
numbers
of stars shown in these plots are $\sim$82\% of the values listed in
Figure~\ref{tab:truth} because here we only count those stars with positions
in $[-1, 1] \times [-1, 1]$.
\label{fig:sex_N}}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
\end{center}
\caption{Inference of the luminosity function parameters using the $0.5\sigma$
catalog from \sex. Note that the inference appears to rule out regions around
the true input values. This occurs because the uncertainties in the existence
and fluxes of the faint stars are not taken into account. The true input values are plotted as filled squares.
\label{fig:sex_inference}}
\end{figure}

@@

We note that there are many limitations to the model presented in this paper.
In principle, our model should be a model of the physical state of the universe,
and not a simple model where the only stellar properties are a 2-D position and
a flux. Another limitation is that we have not considered multi-epoch or multi-band
imaging. In the former case, PSF variations and stellar motions may be relevant
\citep{lang}, and in the latter, a model for the spectral energy distributions
of the stars will need to be considered.

@@

\section{Acknowledgements}
It is a pleasure to thank
  Jonathan Goodman (NYU),
  Fengji Hou (NYU),
  Dustin Lang (CMU),
  Geraint Lewis (Sydney), and
  Phil Marshall (Oxford)
for their comments and discussions.
BJB would like to thank Tommaso Treu (UCSB) for his support, and
Wayne Stewart and Arden Miller (Auckland) for their encouragement and advice.
DFM and DWH were partially supported by
   the US National Science Foundation (grant IIS-1124794) and
   the US National Aeronautics and Space Administration (grant NNX12AI50G).

@@

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcl}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcrl}
0, & & x < h_1 & \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 & \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
Two conditions will be used to determine the normalizers $Z_1$ and $Z_2$.
Firstly, the probability density function (PDF) should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}
The second condition is that the total probability must be 1:
\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}
\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}.
\end{eqnarray}
The cumulative distribution (CDF) is a useful property of a probability
distribution and is given by the antiderivative of the PDF:
\begin{eqnarray}
P(X \leq x) = F(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The inverse of the CDF is also useful and is given by:
\begin{eqnarray}
F^{-1}(u) &=&
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1.
\end{array}
\right.
\end{eqnarray}
An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
\begin{figure}[ht!][ht!]
\begin{center}
\includegraphics[width=\textwidth]{Figures/broken.eps}
\caption{A broken power-law distribution. The parameter values for this
particular PDF were
$\{h_1, h_2, \alpha_1, \alpha_2\} = \{0.3, 0.6, 1.1, 2\}$, i.e. the same
parameter values used to make the simulated data.
\label{fig:powerlaw}}
\end{center}
\end{figure}
